---
title: "4-exploration"
author: "Ashley Clarke"
date: "12/13/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Need to use the following `R` packages and functions:
```{r, message = FALSE}
# load libraries
library(kableExtra)                     # for printing tables
library(cowplot)                        # for side by side plots
library(lubridate)                      # for dealing with dates
library(maps)                           # for creating maps
library(tidyverse)
library(ggcorrplot)                     # for correlation plots
```

```{r read-data, echo=FALSE}
# read in the cleaned data
setwd("~/Desktop/STAT471/unemployment-project")
econ_train = read_tsv(file = "data/clean/econ_train.tsv")
econ_test = read_tsv(file = "data/clean/econ_test.tsv")
```

```{r }
test_obs = nrow(econ_train)
train_obs = nrow(econ_test)
total_obs = test_obs + train_obs
num_features = ncol(econ_train)
```
#The test data set has 585 observations, while the train data set has 147 observations. The data set represents 732 different months of data from June 1960 to December 2020. There are 217 features 

# Variation in response variable 
```{r distribution}
histogram_unemploy = econ_train %>%
  ggplot(aes(x = UNRATE)) +
  geom_histogram(bins = 15, fill = "grey", col =  "black") +
  labs(x = "Unemployment Rate (%)", 
       y = "Count") +
  # add vertical line at the median value for unemployment rate
  geom_vline(xintercept = median(econ_train$UNRATE), color = "#f04546") + 
  theme_bw()

# save the plot
ggsave(filename = "~/Desktop/STAT471/unemployment-project/results/histogram_unemploy.png", 
       plot = histogram_unemploy, 
       device = "png", 
       width = 8, 
       height = 5)
```

```{r top-5-years}
# which 10 years have the highest unemployment rates
# do these correspond with recessions? 

top_10_unemploy = econ_train %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarise(UNRATE = mean(UNRATE)) %>% 
  ungroup() %>%
  arrange(desc(UNRATE)) %>%
  head(10)

top_10_unemploy %>% 
  kable(format = "latex", row.names = NA, 
        booktabs = TRUE, digits = 1) %>%
  kable_styling(position = "center") %>%
  save_kable(file =
     "~/Desktop/STAT471/unemployment-project/results/top_10_unemploy.pdf", 
             self_contained = T)
```

# Calculate mean unemployment 
```{r mean-unemployment}
# All data 
mean_all_years = econ_train %>%
  summarise("Time" = "1960-2020", "Unemployment Rate" = mean(UNRATE))

# After 2010
mean_last_10years = econ_train %>% 
  filter(date > as.Date("2010-01-01")) %>%
  summarise("Time" = "2010-2020", "Unemployment Rate" = mean(UNRATE))

mean_unemploy = rbind(mean_all_years, mean_last_10years) 

mean_unemploy  %>% 
  kable(format = "latex", row.names = NA, 
        booktabs = TRUE, digits = 2) %>%
  kable_styling(position = "center") %>%
  save_kable(file =
     "~/Desktop/STAT471/unemployment-project/results/mean-unemployment-chart.pdf", 
             self_contained = T)
```


#Plot unemployment by year 
```{r all-year-comparison-plot}
# set colors to use for graphs 
cols <- c("Total U.S."= "#3591d1","Mean Unemployment (1972-2021)"= "#f04546")

# plot both unemployment rates for all years
all_year_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x=date,y=UNRATE, colour = "Total U.S.")) +
  geom_hline(yintercept = median(econ_train$UNRATE), 
             colour = "#f04546", linetype='dashed') +
  ggtitle("Total U.S. Unemployment (1960-2020)") + 
  theme_bw() +
  scale_colour_manual(name="Metric",values=cols)  +
  labs(
    x = "Date",
    y = "Unemployment Rate"
  )
# save the plot
ggsave(filename = "~/Desktop/STAT471/unemployment-project/results/all-year-comparison-plot.png", 
       plot = all_year_plot, 
       device = "png", 
       width = 8, 
       height = 5)
```

# Consider Phillips Curve - plot Unemployment vs. Inflation

Does the Phillips Curve hold true for either party?
```{r phillips-curve-US}
# Plot phillips curve by year 
phillips_curve_US = econ_train %>% 
  summarise(year = year(date), UNRATE, FPCPITOTLZGUSA) %>% 
  group_by(year) %>% 
  summarise(unemployment_rate = mean(UNRATE), 
            inflation = mean(FPCPITOTLZGUSA)) %>% ungroup() %>%
  ggplot() + 
  geom_line(mapping = aes(x=unemployment_rate,y=inflation), color = "#3591d1") +
  ggtitle("Inflation vs. Unemployment Rate") + 
  theme_bw() +
  labs(
    x = "Unemployment Rate",
    y = "Inflation"
  )

# save the plot
ggsave(filename = 
         "~/Desktop/STAT471/unemployment-project/results/phillips-curve-us-plot.png", 
       plot = phillips_curve_US, 
       device = "png", 
       width = 8, 
       height = 5)
```

# Plot unemployment rate vs. federal funds rate
```{r federal-funds-plot}
# plot both unemployment rates  
fed_funds_plot = econ_train %>% filter(date > as.Date("2000-01-01")) %>% 
  ggplot() + 
  geom_line(mapping = aes(x= FEDFUNDS,y=UNRATE), color = "#3591d1") +
  ggtitle("Unemployment Rate vs. Federal Funds Rate (2000-2020)") + 
  theme_bw() +
  labs(
    x = "Federal Funds Rate",
    y = "Unemployment Rate"
  )
# save the plot
ggsave(filename = 
         "~/Desktop/STAT471/unemployment-project/results/fed-funds-plot.png", 
       plot = fed_funds_plot, 
       device = "png", 
       width = 8, 
       height = 5)
```
Look at variables related gto manufacturing, mining/logging, construction, information

```{r employees}
# examine unemployment by thousands of employees 

# logging plot
employees_logging_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= CES1011330001, y = UNRATE), color = "#3591d1") +
  ggtitle(" Logging") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Logging Employees (thousands)",
    y = "Unemployment Rate"
  )

# shipping and boating 
employees_shipping_boating_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= CES3133660001, y = UNRATE), color = "#3591d1") +
  ggtitle("Shipping and Boating") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Shipping and Boating Employees (thousands)",
    y = "Unemployment Rate"
  )

# information plot
employees_information_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= CEU5000000001, y = UNRATE), color = "#3591d1") +
  ggtitle("Information") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Information Employees (thousands)",
    y = "Unemployment Rate"
  )

# information plot after 2000, since likely meaningless before then
employees_information_plot_after_2000 = econ_train %>% 
  filter(date > as.Date("2000-01-01")) %>%
  ggplot() + 
  geom_line(mapping = aes(x= CEU5000000001, y = UNRATE), color = "#3591d1") +
  ggtitle("Information (2000-2020)") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Information Employees (thousands)",
    y = "Unemployment Rate"
  )

# federal plot, suggests numbner might increase during high unemployment 
employees_federal_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= CES9091000001, y = UNRATE), color = "#3591d1") +
  ggtitle("Federal") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Federal Employees (thousands)",
    y = "Unemployment Rate"
  )


employees_plot = plot_grid(employees_logging_plot, employees_shipping_boating_plot, employees_information_plot, employees_information_plot_after_2000, employees_federal_plot, rnow = 3)

# save the plot
ggsave(filename = 
         "~/Desktop/STAT471/unemployment-project/results/employees_plot.png", 
       plot = employees_plot, 
       device = "png", 
       width = 8, 
       height = 5)
```





```{r hours-production}
# examine unemployment by hours of work, issue is still classifed as employed even if hours change

# hours manufacturing plot
hours_production_manufacturing_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= AWHMAN, y = UNRATE), color = "#3591d1") +
  ggtitle("Manufacturing") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Hours Production",
    y = "Unemployment Rate"
  )

# hours mining and logging plot
hours_production_mining_logging_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= CES1000000007, y = UNRATE), color = "#3591d1") +
  ggtitle("Mining/Logging") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Hours Production",
    y = "Unemployment Rate"
  )

# hours construction plot
hours_production_construction_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= CEU2000000007, y = UNRATE), color = "#3591d1") +
  ggtitle("Construction") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Hours Production",
    y = "Unemployment Rate"
  )



hours_production_plot = plot_grid(hours_production_manufacturing_plot, hours_production_mining_logging_plot, hours_production_construction_plot, nrow = 1)

# save the plot
ggsave(filename = 
         "~/Desktop/STAT471/unemployment-project/results/hours_production_plot.png", 
       plot = hours_production_plot, 
       device = "png", 
       width = 12, 
       height = 4)
```


```{r federal_aid}
#Government social benefits: to persons: Federal: Benefits from social insurance funds: Unemployment insurance, as increases so should 
social_benefits_plot = econ_train %>% 
  filter(A1589C1A027NBEA <mean(econ_train$A1589C1A027NBEA)*5) %>% #remove a crazy outliar 
  ggplot() + 
  geom_line(mapping = aes(x= A1589C1A027NBEA, y = UNRATE), color = "#3591d1") +
  ggtitle("Social Benefits to Persons") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Government social benefits: to person",
    y = "Unemployment Rate"
  )

#Net government saving: Imputations
net_gov_savings_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= A2010C1A027NBEA, y = UNRATE), color = "#3591d1") +
  ggtitle("Net government saving") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Net government saving",
    y = "Unemployment Rate"
  )

#Real Consumption of Fixed Capital: Private
consumption_fixed_capital_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= A024RL1A225NBEA, y = UNRATE), color = "#3591d1") +
  ggtitle("Real Consumption of Fixed Capital: Private") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Real Consumption of Fixed Capital: Private",
    y = "Unemployment Rate"
  )

#Real Disposable Personal Income
disposable_personal_income_plot = econ_train %>% 
  ggplot() + 
  geom_line(mapping = aes(x= A067RL1A156NBEA, y = UNRATE), color = "#3591d1") +
  ggtitle("Real Disposable Personal Income") + 
  theme_bw() +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  labs(
    x = "Real Disposable Personal Income",
    y = "Unemployment Rate"
  )


gdp_plots = plot_grid(social_benefits_plot, net_gov_savings_plot, consumption_fixed_capital_plot, disposable_personal_income_plot, nrow = 2)

# save the plot
ggsave(filename = 
         "~/Desktop/STAT471/unemployment-project/results/gdp_plots.png", 
       plot = gdp_plots, 
       device = "png", 
       width = 10, 
       height = 6)

```
